Evidence for a Third Color Parameter Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2018-05-30 submit/response.txt

While much of the narrative within the paper is compelling, it is currently unverifiable. Despite the SNFactory project starting 2 decades ago, that most of the data for this paper were taken >1 decade ago, and the data being used in many publications, the spectrophotometric series critical for this analysis (except for perhaps 3 SNe: SNe 2006D, 2011fe, 2012cg) have never been published. there isn't even a table listing the SNe used.

It is impossible to verify any claims without making these data public, and
I cannot accept this paper until the data are released publicly.

Response 1:

For quick reference we have added a table containing a list of the supernovae and their measurements used in this analysis and a second table listing the supernovae and values of their output parameters.  For completeness, the input data and output Monte Carlo chains are made available in a tar archive on a persistent repository (TODO).  The publicly available spectrophotometric series can be used to check how our pipeline transformation the series into the measures used in this article.

It is unfortunate that the spectrophotometric series of all the supernovae used in this analysis are not yet public.  Nevertheless, there is precedent in astrophysics for interesting results that could stimulate public interest and activity (e.g. verification using independent data sets) to be published before all antecedent data products are made public. This article fits that category.

Despite the time delay pointing to the contrary, there is a concentrated effort within the SNfactory collaboration on making a public data release.

@@

Beyond the basic critical idea of providing sufficient data to verify the results, a fundamental aspect of scientific inquiry, the authors do not satisfactory describe their sample. They simply state that the sample of 172 SNe have been processed and have sufficient data to provide some parameters. Does this mean that peculiar SNe Ia are included? Was it harder to reduce the data for some SNe than others (highly embedded SNe)? Presumably there are different selection effects, but none are listed. Without this basic information, it is impossible to fully evaluate this work.

The authors do not discuss if their sample is representative of the broader SN Ia sample or the cosmology samples. In fact, the demographics and parameters are not discussed at all, deferring to Chotard et al. Before any referee can properly review this paper (and in addition to the data being released publicly), Chotard et al. must be submitted to a journal and provided to the referee of this paper.

Response 2:

We have added paragraphs in the Data section that describe the selection of our sample.

A new Figure has been added to show the distributions of the observables in our sample, and the distributions the light—curve observables of the Pantheon sample as a point of reference.

More importantly, we add a discussion on why sample selection is not important for the scope of this analysis.  To summarize, this analysis seeks to detect and model diversity within the Type Ia family.  Diversity found in a biased subsample remains as diversity of the whole family.  While sample selection is critical in the standardization and application of the model across different subsamples, that isn’t the focus of the current article.

@@

Comment 3:

1. It would be useful to show the covariance matrices for all models and discuss how necessary each measurement is. For instance, the Si EW and x1 are highly correlated, and perhaps only one measurement is required.

Response 3:

Note that the correlation/covariance between a subset of parameters are already in the article.  For example the correlation matrix of the per-supernova parameters is given and there is a figure of their distribution.  However, there are too many parameters to make providing their covariance matrix within the article unwieldy. Instead we provide the Monte Carlo chains in the tar archive from which the covariance of any possible parameter pair can be determined.

We add, in Section 5.2.6. the determination of how much dispersion remains in an individual parameter assuming that a second (correlated) parameter is perfectly measured.

@@

Comment 4:

2. Although there is some discussion about the host mass step, it seems like the authors should have a model with this parameter. That is, the mass step is already used for cosmological analyses and it seems prudent to determine if the new parameter is truly a new color parameter or if it is simply the mass step repackaged.

Response 4:

The purpose of this article is to model supernova colors based on properties directly linked to the supernova.  We are then interested in whether the mass-step is a repackaging of our newly-discovered parameter.   The model used in this article is not really suited to capture the mass-step effect, as it does not account for absolute magnitudes.

Comment 5:

@@

3. Section 5.2.9 says that the correlation between Delta and the other parameters is shown in Eq. 20. But Delta isn't included there.

Response 5:

The text in 5.2.9. is a remnant of an old version of the article.  The offending text is now replaced with a reminder of how correlations between Delta and the other parameters are handled in the analysis.

### 2018-05-30 submit/fix.tex
version for VC

@@

\added{
SNfactory observed supernovae in the redshift range $z \lesssim 0.09$, with a preference for $0.03 < z < 0.08$,
whose first SNIFS observation was projected to be $\lesssim -5$ days before $B$-band peak brightness.
Telescope availability and observing conditions dictated which supernovae eventually would accumulate multiple observations 
over the course of their light curves that would give constraining light curve (e.g.\ SALT2) fits.  At an occurence of $< 10\%$ individual spectral extractions 
}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{perobject_input.pdf}
   \caption{Distributions of the central values of the inputs  $EW_{\mathrm{Ca}}$, $EW_{\mathrm{Si}}$,
$\lambda_{\mathrm{Si}}$ , $x_1$, and $C$ of the sample used in this article, and
the $x_1$ and $C$ distributions of the $0.03<z< 0.08$ Pantheon sample,
   \label{input:fig}}
\end{figure}
}

@@

\subsection{Model}
We assume 
that  peak
underlying
 ${\hat{U}}{\hat{B}}\hat{V}\hat{R}\hat{I}$ magnitudes of a supernova  with effects from  astrophysical foregrounds are linearly dependent
on its
 equivalent widths of the Ca~II H\&K and Si~II~$\lambda$4141 spectral features
$EW_{\mathrm{Ca}}$ and $EW_{\mathrm{Si}}$,
and the wavelength of the minimum of 
the Si~II~$\lambda6355$ feature $\lambda_{\mathrm{Si}}$
around $B$-band peak brightness:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2005ApJ...623.1011B, 2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F,2012AJ....143..126B}.
The explicit omission of light-curve shape in our model is compensated by its proxy,
$EW_{\mathrm{Si}}$, at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
The underlying magnitudes are also linearly dependent on the per-supernova latent parameters $g_0$ and $g_1$.
Unlike
the spectral parameters ($EW_{\mathrm{Ca}}$, $EW_{\mathrm{Si}}$ and $\lambda_{\mathrm{Si}}$),  $g_0$ and $g_1$ are not directly associated
with observables but rather are inferred as part of the analysis.
A grey magnitude offset, $\Delta$, is included for each supernova
to capture 
potential
band-independent intrinsic dispersion, while also absorbing peculiar-velocity errors introduced when converting
fluxes to luminosities.
With this grey offset the model standardizes colors, not absolute magnitude.
The underlying magnitudes are
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta \vec{1} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{\mathrm{Ca}} + \beta_{\hat{U}} EW_{\mathrm{Si}} + \eta_{\hat{U}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{\mathrm{Ca}} + \beta_{\hat{B}} EW_{\mathrm{Si}} + \eta_{\hat{B}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{\mathrm{Ca}} + \beta_{\hat{V}} EW_{\mathrm{Si}} + \eta_{\hat{V}} \lambda_{\mathrm{Si}} +\gamma^0_{{\hat{V}}} g_0 +\gamma^1_{{\hat{V}}} g_1 \\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{\mathrm{Ca}} + \beta_{\hat{R}} EW_{\mathrm{Si}} + \eta_{\hat{R}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{R}}} g_0 +\gamma^1_{{\hat{R}}} g_1 \\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{\mathrm{Ca}} + \beta_{\hat{I}} EW_{\mathrm{Si}}+ \eta_{\hat{I}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{I}}} g_0 +\gamma^1_{{\hat{I}}} g_1
\end{pmatrix}.
\label{ewsiv:eqn}
\end{equation}

$\vec{1}$ is the Identity vector.
The global parameter vectors\footnote{Global parameters act on each of the five bands: a single component is written with a subscript for the corresponding band, e.g.\  $c_{\hat{U}}$; 
the combination of all components are written as a vector, e.g.\ $\vec{c}=(c_{\hat{U}}, c_{\hat{B}}, c_{\hat{V}}, c_{\hat{R}}, c_{\hat{I}}) $.  The
band subscript or arrow
notationally identify global parameters, in contrast with the per-supernova parameters.
} that describe the SN~Ia population
$\vec{c}$  (units of mag);
$\vec{\alpha}$, $\vec{\beta}$,
$\vec{\eta}$ (units of mag~\AA$^{-1}$), $\vec{\gamma}^0$ and $\vec{\gamma}^1$ (units of mag)
 are the intercept
and slopes
of the linear relationships that
relate
per-supernova
parameters with 
underlying magnitudes.
The notation used in this article is not to be associated with those from previous works.

@@

>>>>

\added{The current SNfactory sample consists of 198 SNe~Ia
having fully-processed spectrophotometric lightcurve data, including
observations on at least two photometric nights, final references, and a
host spectroscopic redshift. 
These all have at least 5 spectra while the SN is
active, and pass the quality cuts suggested by \citet{2010A&A...523A...7G}.
More than 80\% of our SNe are from searches where there was no
pre-selection based on host galaxy properties (those
whose names start with ``SNF'', ``LSQ'', or ``PTF'' in Table~\ref{data:tab}).

From that sample, we further require at least one spectrum with $\pm 2.5$ days of B-band peak brightness.
We exclude six SNe~Ia in the
SN~1991T, SN~1991bg and SN~2002cx subclasses, two objects 
SNF20080905-005 and SNNGC7589 described in \citet{leget:thesis} as
being spectroscopically peculiar, and two objects SN2012cu and SNF20061108-001 whose SALT2
fits are peculiar.

For the purposes of this article, it is not important whether the population in our sample is representative
of the underlying distribution, either at low- or high-redshift.  This article considers
underlying parameters that
affect SN~Ia colors; such parameters found in a subsample are still
relevant parameters when dealing with the population as a whole.

There is a fraction of SNfactory-observed supernovae that do have the
observations to satisfy the selection criteria, but are excluded due to some observations
having not completed full processing.  Incomplete processing can occur for data taken during poor
observing conditions, when there are instrumental artifacts, and when the supernova is late in its
light-curve and its signal-to-noise and contrast are low.  The sample-selection exclusion of these supernovae are not related,
at least directly, to the observables used in this analysis.  Our results are insensitive to potential sample-selection
bias due to stochasticity in the measurements.

??
The sample selection is insensitive to the above measurements of the spectral features of the spectra that are
closest to maximum light nor of the  peak magnitudes in synthetic
bands interpolated from multiple spectra.
Therefore the results are insensitive
biased due to stochastic measurement uncertainty.

@@

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{perobject_input.pdf}
   \caption{\added{Distributions of the central values of the inputs  $EW_{\mathrm{Ca}}$, $EW_{\mathrm{Si}}$,
$\lambda_{\mathrm{Si}}$ , $x_1$, and $C$ of the sample used in this article, and
the $x_1$ and $C$ distributions of the $0.03<z< 0.08$ Pantheon sample.}
   \label{input:fig}}
\end{figure}
}

@@

\subsection{Model}
We assume 
that  peak
underlying
 ${\hat{U}}{\hat{B}}\hat{V}\hat{R}\hat{I}$ magnitudes of a supernova  with effects from  astrophysical foregrounds are linearly dependent
on its
 equivalent widths of the Ca~II H\&K and Si~II~$\lambda$4141 spectral features
$EW_{\mathrm{Ca}}$ and $EW_{\mathrm{Si}}$,
and the wavelength of the minimum of 
the Si~II~$\lambda6355$ feature $\lambda_{\mathrm{Si}}$
around $B$-band peak brightness:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2005ApJ...623.1011B, 2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F,2012AJ....143..126B}.
The explicit omission of light-curve shape in our model is compensated by its proxy,
$EW_{\mathrm{Si}}$, at peak brightness
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
The underlying magnitudes are also linearly dependent on the per-supernova latent parameters $g_0$ and $g_1$.
Unlike
the spectral parameters ($EW_{\mathrm{Ca}}$, $EW_{\mathrm{Si}}$ and $\lambda_{\mathrm{Si}}$),  $g_0$ and $g_1$ are not directly associated
with observables but rather are inferred as part of the analysis.
A grey magnitude offset, $\Delta$, is included for each supernova
to capture 
potential
band-independent intrinsic dispersion, while also absorbing peculiar-velocity errors introduced when converting
fluxes to luminosities.
With this grey offset the model standardizes colors, not absolute magnitude.
The underlying magnitudes are
\begin{equation}
\begin{pmatrix}
{\hat{U}}\\{\hat{B}}\\{\hat{V}}\\{\hat{R}}\\{\hat{I}}
\end{pmatrix}
=
\Delta \vec{1} +
\begin{pmatrix}
c_{\hat{U}}+\alpha_{\hat{U}} EW_{\mathrm{Ca}} + \beta_{\hat{U}} EW_{\mathrm{Si}} + \eta_{\hat{U}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{U}}} g_0 +\gamma^1_{{\hat{U}}} g_1 \\
c_{\hat{B}}+\alpha_{\hat{B}} EW_{\mathrm{Ca}} + \beta_{\hat{B}} EW_{\mathrm{Si}} + \eta_{\hat{B}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{B}}} g_0 +\gamma^1_{{\hat{B}}} g_1 \\
c_{\hat{V}}+\alpha_{\hat{V}} EW_{\mathrm{Ca}} + \beta_{\hat{V}} EW_{\mathrm{Si}} + \eta_{\hat{V}} \lambda_{\mathrm{Si}} +\gamma^0_{{\hat{V}}} g_0 +\gamma^1_{{\hat{V}}} g_1 \\
c_{\hat{R}}+\alpha_{\hat{R}} EW_{\mathrm{Ca}} + \beta_{\hat{R}} EW_{\mathrm{Si}} + \eta_{\hat{R}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{R}}} g_0 +\gamma^1_{{\hat{R}}} g_1 \\
c_{\hat{I}}+\alpha_{\hat{I}} EW_{\mathrm{Ca}} + \beta_{\hat{I}} EW_{\mathrm{Si}}+ \eta_{\hat{I}} \lambda_{\mathrm{Si}}  +\gamma^0_{{\hat{I}}} g_0 +\gamma^1_{{\hat{I}}} g_1
\end{pmatrix}.
\label{ewsiv:eqn}
\end{equation}
$\vec{1}$ is the Identity vector.
The global parameter vectors\footnote{Global parameters act on each of the five bands: a single component is written with a subscript for the corresponding band, e.g.\  $c_{\hat{U}}$; 
the combination of all components are written as a vector, e.g.\ $\vec{c}=(c_{\hat{U}}, c_{\hat{B}}, c_{\hat{V}}, c_{\hat{R}}, c_{\hat{I}}) $.  The
band subscript or arrow
notationally identify global parameters, in contrast with the per-supernova parameters.
} that describe the SN~Ia population
$\vec{c}$  (units of mag);
$\vec{\alpha}$, $\vec{\beta}$,
$\vec{\eta}$ (units of mag~\AA$^{-1}$), $\vec{\gamma}^0$ and $\vec{\gamma}^1$ (units of mag)
 are the intercept
and slopes
of the linear relationships that
relate
per-supernova
parameters with 
underlying magnitudes.
The notation used in this article is not to be associated with those from previous works.

@@

\added{This model is applied to our supernova sample and so the ensuing results describe the the SN~Ia population
weighted by our sample selection.}
